\section{Future Work}
This work is focused on the second-order SBP-SAT methods. SBP-SAT methods are known to be high order accuracy, and using higher-order SBP operators will increase arithmetic intensity that will increase the performance of matrix-free GPU kernels compared to SpMV operators.

In addition, the matrix-free kernels presented in this paper are only implemented on a single GPU. 
Although this is enough to solve the problems presented in this paper, it would be more important in the HPC aspect to design matrix-free kernels that can run on multiple GPUs across different nodes.
ParallelStencils.jl is a Julia package that enables large-scale stencil-based computations using built-in modules that utilize MPI for communication.
It's also built on top of KernelAbstractions.jl, a Julia package that targets heterogeneous platforms.
Our code on matrix-free SBP-SAT methods can be developed with ParallelStencils.jl to run on supercomputers built with CPUs/GPUs from different vendors.

In this work, we only apply Richardson iteration as matrix-free smoothers for our multigrid preconditioners. During our research, we observed faster convergence with higher-order Krylov subspace methods as smoother. These methods are also highly suitable for GPU architectures and can be implemented matrix-free.
In future work, we can explore using higher order second-order Richardson methods and Chebyshev iterations as smoothers in multigrid methods to further reduce steps till convergence for CG.