% The acknowledgements section is optional. It can cover at most two pages.
\acknowledge{
	% If don't have an acknowledgment, comment out or delete this whole file
	% To do \dots \\
	My Ph.D. journey would not have been possible without the guidance and help of several individuals. I want to dedicate this part to acknowledge their contribution to this trip.

	First and foremost, I would like to express my deepest gratitude to my advisor, Dr. Brittany A. Erickson. I want to express my deepest gratitude to my advisor, Dr. Brittany A. Erickson. She took me on as her first PhD student during a particularly challenging time in my life, when I was questioning my life choices to come here at the beginning of graduate school. Her unwavering support and guidance have been invaluable, teaching me how to conduct research in applied mathematics and write papers professionally. She gives me the freedom to explore various fields of computer science and applied mathematics while providing technical support in problem formulation and code development whenever I need it. Without her support throughout the years, this thesis would not be possible.
	In addition to her role as an advisor, I also want to express my deepest gratitude for her support as a friend to her students. The past 6 years have been the most challenging period for an international student from China studying here. The mix of trade war, geopolitical conflict, COVID-19, and the post-pandemic economic downturn in the technology sector has made my journey challenging. Whether it's up or down in my life and work, Brittany is always there to celebrate my achievements and share my sadness. I was also fortunate to witness her amazing achievements in her career and family over the past years. I will continue to see Brittany as a role model for inspiration and motivation in my future work and life.

	I am also deeply grateful to my committee members, Dr. Boyana Norris, Dr. Jee Whan Choi, Dr. Richard Tran Mills, and Dr. Leif Karlstrom. Dr. Boyana Norris has been serving as a committee member since my Directed Research Project (DRP). 
	She guided me in various research directions and consistently provided valuable feedback throughout my program. I was also invited to join her group meeting, through which I got opportunities to present my work and meet other members of her group.
	She was also very generous in purchasing a great espresso machine for the common office space that I am sharing with students in her group. 
	It has been the number one motivation for me to come to the office, and I am one of the largest beneficiaries of the espresso machine.
	I want to thank Dr. Jee Whan Choi for his guidance in my research, especially in analyzing the performance of the GPU computing part. Coming from a physics undergraduate, I have little background in computer systems that can support me in conducting research in HPC. I learned about HPC and GPU computing much faster from direct collaboration with Dr. Jee Whan Choi. He contributed so much to help me get my first publication in HPC which this thesis is based on. Without his support, this thesis would not be possible.
	I want to express my gratitude to Dr. Richard Tran Mills for his dedicated work on PETSc and for his assistance when we integrated PETSc into our project. I also appreciate his presence at my dissertation defense and his thorough review of my thesis draft, where he provided valuable constructive feedback.
	I also want to thank Dr. Leif Karlstrom for serving on my committee and for raising many important geophysics-related questions. His insights encouraged me to view my research from a broader perspective and to explore ways to maximize the impact of my current results in my future work.


	I want to extend my gratitude to Dr. Jeremy E. Kozdon for teaching me HPC and GPU computing in Julia and guiding me through my DRP work. His feedback and encouragement throughout my PhD program were invaluable. Additionally, I want to thank Dr. Hank Childs from the CS department for his support and care for my research, as well as his dedication to all graduate students. For my research work, I want to thank Dr. Sameer Shende for providing me access to Oregon Advanced Computing Institute for Science and Society (OACISS) to accelerate my research. 
	I also want to express my gratitude to many colleagues in the CS department, including the incredible graduate coordinators Cheri Smith and Nicole Moynahan, who helped me navigate the PhD program, as well as numerous graduate students who have shared this journey with me.

	A special thanks to my family for their unwavering love and support. To my parents, Liugen (Thomas) Chen and Yuexia (Luna) Li, thank you for bringing me to this world and encouraging me to pursue my dreams. 

	This work benefited from access to the University of Oregon high performance computing cluster, Talapas, and Oregon Advanced Computing Institute for Science and Society (OACISS). This work is supported by NSF awards \#2053372 and \#2036980. 
}

